### 2019-02-13 sections/intuition.tex
add todo
some stuff in intuition

The central objective of tasks is to \emph{coordinate collaboration}.
Therefore it has combinators to express ways in which people can work together to reach a common goal.
Tasks can be executed after each other, at the same time, or conditionally.
This motivates the task combinators for steps, parallel combination, and choice.
One can say that tasks are a way to model \emph{workflows}.
% \TOPHAT is a language to specify workflows.

@@

@@

In \TOPHAT, communication with the environment is represented by editors.
A task of the form $\Edit v$, where $v$ is of type $\tau$, can receive any change event with a value of type $\tau$ and additionally the special event $\Empty$ that empties the editor.
There is no explicit communication of output.
% However, the environment can always inspect the current value of an editor.

@@

This example demonstrates external communication and choice.
The following \CSP program models a vending machine that dispenses a small biscuit for one coin and a large biscuit for two coins.
\begin{equation*}
  \textit{coin} \to (\textit{small} \to \text{STOP} \mid \textit{coin} \to \textit{large} \to \text{STOP})
\end{equation*}
% An equivalent reactive system in \ESTEREL could be defined as\fixme{validate}
% \begin{equation*}
  % V_\text{\ESTEREL} = \text{await } \textit{coin}; ( \text{await } \textit{button}; \text{emit } \textit{small} \parallel \text{await } \textit{coin}; \text{await } \textit{button}; \text{emit } \textit{large})
% \end{equation*}

@@

Because every task has an observable value of a specified type,
it is always possible to generate a (graphical) user interface from a \TOPHAT specification.
%NOTE: depends on: editors
The structure of the specification prescribes the structure of the \UI and possible input events.
Editors form the leaves of such trees.

@@

The decision to move on to the next task is not performed by the task itself.
It is based on a calculation to determine the next tasks.
The step combinators in $t \Then e$ and $t \Next e$ pass the value of $t$ to the continuation $e$.
They both act like $t$ as long as the step is guarded.
Once it becomes unguarded, the task continues as $e\ v$, where $v$ is the observed value of $t$.
% The continuation $e$ is a function which calculates the next task or $\Fail$, signalling no continuation is present.
The task $t \Next e$ requires an input event $\Continue$ in addition to the step being unguarded. % in order to step.

@@

\fixme{?}
This introduces the possibility of diverging computations, where a cyclic dependency between shared data causes internal steps to be taken continuously without ever being able to engage in external communication.

@@

While tasks are focussed on collaboration, they give rise to concepts like communication, concurrency and synchronisation.
Using tasks, however, they are meant to be used on a higher level of abstraction, which allows concise declarative description of patterns of collaboration, without worrying about the implementation details of communication.
By focussing on collaboration, tasks and \TOPHAT lead to high-level specifications closer to real-world workflows which, at the same time, can be used to generate distributive, multi-user applications to support such workflows.

>>>>

The central objective of tasks is to \emph{coordinate collaboration}.
Therefore \TOPHAT has combinators to express ways in which people can work together to reach a common goal.
Tasks can be executed after each other, at the same time, or conditionally.
This motivates the task combinators for steps, parallel combination, and choice.
One can say that tasks are a way to model \emph{workflows}.
% \TOPHAT is a language to specify workflows.

@@

\fixme{Add example about higher order tasks}

@@

In \TOPHAT, communication with the environment is represented by editors.
A task of the form $\Edit v$, where $v$ is of type $\tau$, can receive any change event with a value of type $\tau$ and additionally the special event $\Empty$ that empties the editor.
There are no output events, instead the environment can always inspect a task's value.
% However, the environment can always inspect the current value of an editor.

@@

This example demonstrates external communication and choice.
The following \CSP program models a vending machine that dispenses a small biscuit for one coin and a large biscuit for two coins. \todo{remove CSP?}
\begin{equation*}
  \textit{coin} \to (\textit{small} \to \text{STOP} \mid \textit{coin} \to \textit{large} \to \text{STOP})
\end{equation*}
% An equivalent reactive system in \ESTEREL could be defined as\fixme{validate}
% \begin{equation*}
  % V_\text{\ESTEREL} = \text{await } \textit{coin}; ( \text{await } \textit{button}; \text{emit } \textit{small} \parallel \text{await } \textit{coin}; \text{await } \textit{button}; \text{emit } \textit{large})
% \end{equation*}

@@

Because every task has an observable value of a specified type,
it is always possible to generate a (graphical) user interface from a \TOPHAT specification.
%NOTE: depends on: editors
The structure of the specification prescribes the structure of the \UI and possible input events.
Editors form the leaves of such trees \todo{which trees?}.

@@

The decision to move on to the next task is not taken by the task itself.
It is based on a calculation to determine the next tasks.
The step combinators in $t \Then e$ and $t \Next e$ pass the value of $t$ to the continuation $e$.
They both act like $t$ as long as the step is guarded.
Once it becomes unguarded, the task continues as $e\ v$, where $v$ is the observed value of $t$.
% The continuation $e$ is a function which calculates the next task or $\Fail$, signalling no continuation is present.
The task $t \Next e$ requires an input event $\Continue$ in addition to the step being unguarded. % in order to step.

@@

@@

The focus on collaboration gives rise to concepts like communication, concurrency and synchronisation.
Using tasks, however, they are meant to be used on a higher level of abstraction, which allows concise declarative description of patterns of collaboration, without worrying about the implementation details of communication.
By focussing on collaboration, tasks and \TOPHAT lead to high-level specifications closer to real-world workflows which, at the same time, can be used to generate distributive, multi-user applications to support such workflows.
